### 2016-05-04 MS.tex

\section{\label{Known}Known metric solutions to the parallel Ricci equations}

In this Appendix, we will present a brief compendium of known (Riemannian) metric solutions of Eq.~\eqref{SimpleEOM} in four dimensions. The results presented in this Appendix is not original, and can be found in Refs.~\cite{gray1978einstein,derdzinski1980classification,derdzinski1982compact,Derdzinski:1985,Besse}.

\paragraph{Einstein spaces.---} The simplest example of a manifold with parallel Ricci is a Riemannian Einstein manifols. This example was explained on the main text.

\paragraph{Riemannian products.---} The Riemamian product of manifolds with harmonic curvature.

\paragraph{Conformally flat manifolds.---} These are solutions in four dimensions if their scalar curvature is constant.

\paragraph{Warped products.---} Compact warped products of the form \mbox{$(S^1 \times \bar{M}, \de{t}^2 + f(t) \de{s}^2(\bar{g}) )$,} where $(\bar{M}, \bar{g})$ is an Einstein manifold with positive scalar curvature, $\bar{R} > 0$, and $f$ is a positive function on $S^1$ satisfying the differential equation
\begin{equation}
  \ddot{f} - \frac{1}{3} \bar{R} = c f,
\end{equation}
for a negative constant $c$.

{Twisted} warped products of the form $(\R \times \bar{M} ) / \Z$ have parallel Ricci, if the $\Z$-action on the product metric involves an isometry of $\bar{g}$.

Compact warped products of the form \mbox{$(M_1 \times M_2, f^{2} \cdot (g_1 \times g_2) )$,} where $M_i$ are two two-dimensional manifolds, $M_1$ has constant Gau\ss{}ian curvature, $K_1 < 0$, and $M_2$ is an Einstein manifold with scalar curvature $R_2 = - 2 K_1$. The function $f: M_1 \to \R^+$ is a $C^\infty$ solution of 
\begin{equation*}
  \Lap f = c \, f^3 \text{ for } c > 0.
\end{equation*}

Many examples and explicit conditions of Riemannian metrics with harmonic curvature, and conditions for them to have parallel Ricci, can be found in Ref.~\cite{derdzinski1988riemannian}.

Although the above are examples of manifolds with parallel Ricci within the context of ``Riemannian'' geometries, examples of connnections with parallel Ricci can be found in symplectic geometries, see Ref.~\cite{cahen2000symplectic}.

\section{\label{Schw}Explicit example: Schwarzschild ansatz, and Birkhoff theorem}

%% As a disclaimer\hl{What is this... Should this be dropped?}, we clarify that the example below is a pseudo-Riemannian manifold.

%% \subsubsection{Schwarzschild metric}

As pointed out, all vacuum solutions of Einstein's equations ---with and without cosmological constant--- satisfy the parallel Ricci condition trivially~\cite{bourguignon1981varietes}. Nonetheless, an important result in General Relativity is the stability of the Schwarzschild metric, known as the Birkhoff theorem~\cite{Jebsen1921,Birkhoff1923,Alexandrow1923,Eisland1925}, which implies that the gravitational collapse of a spherically symmetric astrophysical object cannot emit gravitational radiation.

In this section we show that the simplest ansatz of a static and spherically symmetric metric, which solves the Eq.~\eqref{SimpleEOM} is necessarily the $\Lambda$--Schwarzschild metric. This should be taken as a first step toward a Birkhoff theorem, despite the fact that a complete proof is far from been achieved.

Consider first the simplest metric ansatz, whose line element is
\begin{equation}
  \de{s}^2 = - f(t,r) \de{t}^2 + \frac{ \de{r}^2 }{ f(t,r) } + r^2 \de{\Omega}^2.
  \label{simpleSchw}
\end{equation}
\begin{widetext}
  The nontrivial field equation for the ansatz~\eqref{simpleSchw} are
  \begin{dmath}
    \nabla_{[\mu} R_{\nu]\lambda} = \frac{1}{2 \, r^{2} f^{3}} \Bigg( r^{2} f^{4} \frac{\partial^3}{\partial r^{3}}f + 2 \, r f^{4} \frac{\partial^2}{\partial r^{2}}f + r^{2} f^{2} \frac{\partial^3}{\partial t^{2}\partial r^{}}f - 4 \, r^{2} f \frac{\partial}{\partial t^{}}f \frac{\partial^2}{\partial t^{}\partial r^{}}f - 2 \, f^{4} \frac{\partial}{\partial r^{}}f + 2 \, {\left(3 \, r^{2} \frac{\partial}{\partial r^{}}f - 2 \, r f\right)} \frac{\partial}{\partial t^{}}f^{2} - 2 \, {\left(r^{2} f \frac{\partial}{\partial r^{}}f - r f^{2}\right)} \frac{\partial^2}{\partial t^{2}}f \Bigg) \,
    \delta_{\mu\nu}^{10} \delta_\lambda^0
    +
    \frac{1}{2 \, r^{2} f^{5}} \Bigg( r^{2} f^{4} \frac{\partial^3}{\partial t^{}\partial r^{2}}f + 2 \, f^{4} \frac{\partial}{\partial t^{}}f + 6 \, r^{2} \frac{\partial}{\partial t^{}}f^{3} - 6 \, r^{2} f \frac{\partial}{\partial t^{}}f \frac{\partial^2}{\partial t^{2}}f + r^{2} f^{2} \frac{\partial^3}{\partial t^{3}}f \Bigg) \,
    \delta_{\mu\nu}^{10} \delta_\lambda^1
    +
    r \frac{\partial^2}{\partial t^{}\partial r^{}}f \,
    \Big( \delta_{\mu\nu}^{20} \delta_\lambda^2 + \sin^2(\theta) \delta_{\mu\nu}^{30} \delta_\lambda^3 \Big)
    +
    \frac{1}{2 \, r f^{3}} \Bigg(r^{2} f^{3} \frac{\partial^2}{\partial r^{2}}f - 2 \, f^{4} + 2 \, r^{2} \frac{\partial}{\partial t^{}}f^{2} - r^{2} f \frac{\partial^2}{\partial t^{2}}f + 2 \, f^{3} \Bigg) \,
    \Big( \delta_{\mu\nu}^{21} \delta_\lambda^2 + \sin^2(\theta) \delta_{\mu\nu}^{31} \delta_\lambda^3 \Big)
    = 0
  \end{dmath}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  %%   \begin{dmath}
  %%   \nabla_{[1} R_{0]0} = \frac{1}{2 \, r^{2} f^{3}} \Bigg( r^{2} f^{4} \frac{\partial^3}{\partial r^{3}}f + 2 \, r f^{4} \frac{\partial^2}{\partial r^{2}}f + r^{2} f^{2} \frac{\partial^3}{\partial t^{2}\partial r^{}}f - 4 \, r^{2} f \frac{\partial}{\partial t^{}}f \frac{\partial^2}{\partial t^{}\partial r^{}}f - 2 \, f^{4} \frac{\partial}{\partial r^{}}f + 2 \, {\left(3 \, r^{2} \frac{\partial}{\partial r^{}}f - 2 \, r f\right)} \frac{\partial}{\partial t^{}}f^{2} - 2 \, {\left(r^{2} f \frac{\partial}{\partial r^{}}f - r f^{2}\right)} \frac{\partial^2}{\partial t^{2}}f \Bigg) ,
  %%   \label{e100}
  %% \end{dmath}
  %% \begin{dmath}
  %%   \nabla_{[1} R_{0]1} = \frac{1}{2 \, r^{2} f^{5}} \Bigg( r^{2} f^{4} \frac{\partial^3}{\partial t^{}\partial r^{2}}f + 2 \, f^{4} \frac{\partial}{\partial t^{}}f + 6 \, r^{2} \frac{\partial}{\partial t^{}}f^{3} - 6 \, r^{2} f \frac{\partial}{\partial t^{}}f \frac{\partial^2}{\partial t^{2}}f + r^{2} f^{2} \frac{\partial^3}{\partial t^{3}}f \Bigg) .
  %%   \label{e101}
  %% \end{dmath}
  %% \begin{dmath}
  %%   \nabla_{[2} R_{0]2} = r \frac{\partial^2}{\partial t^{}\partial r^{}}f,
  %%   \label{e202}
  %% \end{dmath}
  %% \begin{dmath}
  %%   \nabla_{[2} R_{1]2} = \frac{1}{2 \, r f^{3}} \Bigg(r^{2} f^{3} \frac{\partial^2}{\partial r^{2}}f - 2 \, f^{4} + 2 \, r^{2} \frac{\partial}{\partial t^{}}f^{2} - r^{2} f \frac{\partial^2}{\partial t^{2}}f + 2 \, f^{3} \Bigg),
  %%   \label{e212}
  %% \end{dmath}
  %% %% \begin{dmath}
  %% %%   \nabla_{[3} R_{0]3} = r \sin\left({\theta}\right)^{2} \frac{\partial^2}{\partial t^{}\partial r^{}}f,
  %% %%   \label{e303}
  %% %% \end{dmath}
  %% %% and
  %% %% \begin{dmath}
  %% %%   \nabla_{[3} R_{1]3} = \frac{1}{2 \, r f^{3}} \Bigg(r^{2} f^{3} \frac{\partial^2}{\partial r^{2}}f - 2 \, f^{4} + 2 \, r^{2} \frac{\partial}{\partial t^{}}f^{2} - r^{2} f \frac{\partial^2}{\partial t^{2}}f + 2 \, f^{3}\Bigg) \sin\left({\theta}\right)^{2}.
  %% %%   \label{e313}
  %% %% \end{dmath}
  %% \begin{align}
  %%   \nabla_{[3} R_{0]3} &= \sin\left({\theta}\right)^{2} \nabla_{[2} R_{0]2}, &
  %%   \nabla_{[3} R_{1]3} &= \sin\left({\theta}\right)^{2} \nabla_{[2} R_{1]2}.
  %% \end{align}
\end{widetext}
Even though four out of the six nontrivial field equations are independent, the system can be solve uniquely. First, the component along the $(\mu\nu\lambda) = (2,0,2)$ yields $f = T(t) + R(r)$. Then, one notices that the equation along $(\mu\nu\lambda) = (1,0,1)$ contains a term depending solely on $t$, which implies that $\dot{T}$ vanishes, i.e., $T$ is a constant. From here, the equation for $(\mu\nu\lambda) = (2,1,2)$ gets simple enough to be solved and yields, \mbox{$R(r) = 1 - T + \frac{\alpha}{r} + \beta r^2$} or equivalently
\begin{equation}
  f(r) = 1 + \frac{\alpha}{r} + \beta r^2,
  \label{simplef}
\end{equation}
with $\alpha$ and $\beta$ integration constants. Finally, it can be checked that albeit the equation along the components $(\mu,\nu,\lambda) = (1,0,0)$ was not used, it is satisfied by Eq.~\eqref{simplef}.

The above shows that the simplest time-dependent generalization of Schwarzschild metric is not a solution for the field equations of the considered affine model. Unless, the spacetime metric is static and exactly that for a Schwarzschild exterior solution, with or without cosmological constant.

However, if one tries a metric ansatz with two different factors, say
\begin{equation}
  \de{s}^2 = - A(t,r) \de{t}^2 + \frac{ \de{r}^2 }{ B(t,r) } + r^2 \de{\Omega}^2,
  \label{ABmetric}
\end{equation}
the field equations are so complicated, that although hints on the solubility were found ---and yield to Eq.~\eqref{simplef} for both $A$ and $B$---, the uniqueness of the solution is far for been proved. The field equations for the metric~\eqref{ABmetric} are %%shown in Appendix~\ref{app:eqAB}.
\begin{widetext}
  \begin{dmath}
    \nabla_{[\mu} R_{\nu]\lambda} = \frac{1}{4 \, r^{2} B^{3} A^{2}} \Bigg( 2 \, r^{2} B^{4} \frac{\partial}{\partial r^{}}A^{3} + 2 \, r^{2} B^{4} A^{2} \frac{\partial^3}{\partial r^{3}}A + 2 \, r^{2} B^{2} A^{2} \frac{\partial^3}{\partial t^{2}\partial r^{}}B - 6 \, r^{2} B A^{2} \frac{\partial}{\partial t^{}}B \frac{\partial^2}{\partial t^{}\partial r^{}}B - r^{2} B^{2} A \frac{\partial}{\partial t^{}}B \frac{\partial^2}{\partial t^{}\partial r^{}}A + 6 \, {\left(r^{2} A^{2} \frac{\partial}{\partial r^{}}B - r B A^{2}\right)} \frac{\partial}{\partial t^{}}B^{2} - 2 \, {\left(r^{2} B^{3} A \frac{\partial}{\partial r^{}}B + r B^{4} A\right)} \frac{\partial}{\partial r^{}}A^{2} - 2 \, {\left(r^{2} B A^{2} \frac{\partial}{\partial r^{}}B - 2 \, r B^{2} A^{2}\right)} \frac{\partial^2}{\partial t^{2}}B - {\left(r^{2} B^{2} A \frac{\partial^2}{\partial t^{}\partial r^{}}B - 2 \, r^{2} B^{2} \frac{\partial}{\partial t^{}}B \frac{\partial}{\partial r^{}}A - {\left(r^{2} B A \frac{\partial}{\partial r^{}}B - 2 \, r B^{2} A\right)} \frac{\partial}{\partial t^{}}B\right)} \frac{\partial}{\partial t^{}}A + {\left(r^{2} B^{3} A^{2} \frac{\partial^2}{\partial r^{2}}B + 2 \, r B^{3} A^{2} \frac{\partial}{\partial r^{}}B - 4 \, B^{4} A^{2} + 3 \, r^{2} B A \frac{\partial}{\partial t^{}}B^{2} - 2 \, r^{2} B^{2} A \frac{\partial^2}{\partial t^{2}}B\right)} \frac{\partial}{\partial r^{}}A + {\left(3 \, r^{2} B^{3} A^{2} \frac{\partial}{\partial r^{}}B - 4 \, r^{2} B^{4} A \frac{\partial}{\partial r^{}}A + 4 \, r B^{4} A^{2}\right)} \frac{\partial^2}{\partial r^{2}}A \Bigg) \,
    \delta_{\mu\nu}^{10} \delta_\lambda^0
    +
    \frac{1}{4 \, r^{2} B^{4} A^{3}} \Bigg( 2 \, r^{2} B^{4} A^{2} \frac{\partial^3}{\partial t^{}\partial r^{2}}A + r^{2} B^{3} A^{2} \frac{\partial^2}{\partial t^{}\partial r^{}}B \frac{\partial}{\partial r^{}}A - r^{2} B^{3} A \frac{\partial}{\partial t^{}}B \frac{\partial}{\partial r^{}}A^{2} + 2 \, r^{2} B^{3} A^{2} \frac{\partial}{\partial t^{}}B \frac{\partial^2}{\partial r^{2}}A + 4 \, B^{3} A^{3} \frac{\partial}{\partial t^{}}B + 6 \, r^{2} A^{2} \frac{\partial}{\partial t^{}}B^{3} - 8 \, r^{2} B A^{2} \frac{\partial}{\partial t^{}}B \frac{\partial^2}{\partial t^{2}}B + 2 \, r^{2} B^{2} A^{2} \frac{\partial^3}{\partial t^{3}}B + 2 \, r^{2} B^{2} \frac{\partial}{\partial t^{}}B \frac{\partial}{\partial t^{}}A^{2} - r^{2} B^{2} A \frac{\partial}{\partial t^{}}B \frac{\partial^2}{\partial t^{2}}A - {\left(r^{2} B^{3} A \frac{\partial}{\partial r^{}}B \frac{\partial}{\partial r^{}}A - 2 \, r^{2} B^{4} \frac{\partial}{\partial r^{}}A^{2} + 2 \, r^{2} B^{4} A \frac{\partial^2}{\partial r^{2}}A - 4 \, r^{2} B A \frac{\partial}{\partial t^{}}B^{2} + 3 \, r^{2} B^{2} A \frac{\partial^2}{\partial t^{2}}B\right)} \frac{\partial}{\partial t^{}}A + {\left(r^{2} B^{3} A^{2} \frac{\partial}{\partial r^{}}B - 2 \, r^{2} B^{4} A \frac{\partial}{\partial r^{}}A\right)} \frac{\partial^2}{\partial t^{}\partial r^{}}A \Bigg) \,
    \delta_{\mu\nu}^{10} \delta_\lambda^1
    +
    \frac{1}{2 \, A^{2}} \Bigg( r A^{2} \frac{\partial^2}{\partial t^{}\partial r^{}}B + r B A \frac{\partial^2}{\partial t^{}\partial r^{}}A + r A \frac{\partial}{\partial t^{}}B \frac{\partial}{\partial r^{}}A - r B \frac{\partial}{\partial t^{}}A \frac{\partial}{\partial r^{}}A \Bigg) \,
    \Big( \delta_{\mu\nu}^{20} \delta_\lambda^2 + \sin^2(\theta) \delta_{\mu\nu}^{30} \delta_\lambda^3 \Big)
    +
    \frac{1}{4 \, r B^{2} A^{2}} \Bigg( 2 \, r^{2} B^{2} A^{2} \frac{\partial^2}{\partial r^{2}}B + r^{2} B^{2} A \frac{\partial}{\partial r^{}}B \frac{\partial}{\partial r^{}}A - r^{2} B^{3} \frac{\partial}{\partial r^{}}A^{2} + 3 \, r^{2} A \frac{\partial}{\partial t^{}}B^{2} - 2 \, r^{2} B A \frac{\partial^2}{\partial t^{2}}B + r^{2} B \frac{\partial}{\partial t^{}}B \frac{\partial}{\partial t^{}}A - 4 \, {\left(B^{3} - B^{2}\right)} A^{2} \Bigg) \,
    \Big( \delta_{\mu\nu}^{21} \delta_\lambda^2 + \sin^2(\theta) \delta_{\mu\nu}^{31} \delta_\lambda^3 \Big)
    = 0,
    \label{hugeeq}
  \end{dmath}
\end{widetext}
which are very complex, and ---to our knowledge--- these are not possible to solve in general. Nonetheless, the symmetries in some of the components of these equations seems to favour the relation $A = B$ as solution. In addition, we found a very interesting reference in which a proof of the Birkhoff theorem is given for an Einstein--Cartan theory coupled with the gravitational Yang--Mills~\cite{Ramaswamy:1979zz}, which support the hand-waving symmetry argument above.

Hence, further analysis requires a new strategy. Instead of considering a metric, a connection ansatz is proposed based in the structure of a spherically symmetric metric of the form,
\begin{equation}
  \de{s}^2 = g_{ab}(x) \, \de{x}^a\de{x}^b + e^{2\rho(x)} \, g_{AB}(\vph)  \de{\vph}^A\de{\vph}^B,
  \label{sphere-gen}
\end{equation}
where $x^m$ represent the temporal an radial directions, while $\vph^M$ represent angular coordinates. Notice that the four-dimensional indices have been partitioned as \mbox{$\mu = (a,A)$,} then the general form of the connection is
%~\footnote{For the metric of Eq.~\eqref{sphere-gen}, the factors are \mbox{$M^a(x) = - e^{2\rho(x)} g^{am}\partial_m \rho$} and \mbox{$N_c = \partial_c \rho$.}}
\begin{dmath}
  \Ga^\lambda{}_{\mu\nu} = \delta^\lambda_a \delta^b_\mu \delta^c_\nu \ga^a{}_{bc} + \delta^\lambda_A \delta^B_\mu \delta^C_\nu \ga^A{}_{BC} + \delta^\lambda_a \delta^B_\mu \delta^C_\nu g_{BC} M^a(x) +  \delta^\lambda_A \delta^A_{(\mu} \delta^c_{\nu)} N_{c}(x),
  \label{connSchw}
\end{dmath}
where $\ga^a{}_{bc}$ and $\ga^A{}_{BC}$ are the Levi-Civita connection for the metrics $g_{ab}$ and $g_{AB}$ respectively, and the factors are \mbox{$M^a(x) = - \tfrac{1}{2}  g^{am}\partial_m e^{2\rho(x)}$} and \mbox{$N_c = 2 \partial_c \rho$}.

From the connection ansatz~\eqref{connSchw} the Ricci tensor is
\begin{dmath}
  R_{\mu\nu}
  = \delta_\mu^m \delta_\nu^n \( R_{mn} - 2 \nab{m} N_n - 2 N_m N_n \)
  + \delta_\mu^M \delta_\nu^N \( R_{MN} + g_{MN} \nab{a} M^a \),
  %% + ( \delta_\mu^M \delta_\nu^n + \delta_\mu^n \delta_\nu^M ) \gamma^A{}_{AM} N_n
\end{dmath}
and the equations of motion are
\begin{widetext}
  \begin{dmath}
    \nabla_{[\lambda} R_{\mu]\nu}
    = \delta_\lambda^l \delta_\mu^m \delta_\nu^n  \bigg( \frac{1}{2} g_{n[m} \partial_{l]} \mathcal{R} + 2 \mathcal{R}  g_{n[m} N_{l]}(x) - 2 \nabla_{[l} \Big( N_{m]} N_n \Big) \bigg)
    + \delta_\lambda^{[l} \delta_\mu^{M]} \delta_\nu^N \, g_{MN} \bigg( \nab{l} \nab{a} M^a + \frac{\mathcal{R}}{2} g_{la} M^a - 2 N_l - 2 \nab{a} \Big( M^a N_l \Big) - 2 M^a N_a N_l \bigg).
    \label{OurEq}
  \end{dmath}
\end{widetext}
We have used that $g_{AB}(\varphi)$ is the metric of a two-sphere, and its curvature tensors are well known. In addition, since the metric $g_{ab}(x)$ is two-dimensional too, its curvature tensors are related with the metric by the relation
\begin{dmath}
  R_{ab}{}^c{}_d = \frac{\mathcal{R}}{2} \Big( \delta^c_a g_{bd} - \delta^c_b g_{ad} \Big).
\end{dmath}

As argue  in Refs.~\cite{bergmann1965spherically,hawking1973large},  the solution depends on the nature of the surfaces described by $\rho = \text{const.}$, which in our parametrization translate to the nature of $N_a$. We first consider the condition $g^{ab} N_a N_b > 0 $, which allows to set $\rho$ as a good radial coordinate. Therefore, the general metric ansatz $g_{ab}$ is diagonal, and the radial direction would be $\rho$, i.e.,
\begin{equation}
  g_{ab} = - \alpha^2(t,\rho) \de{t}_a \otimes \de{t}_b + \beta^2(t,\rho) \de{\rho}_a \otimes \de{\rho}_b.
  \label{metr2d}
\end{equation}
With this ansatz, the nontrivial components of Eq.~\eqref{OurEq} are
\begin{align}
  \beta^2 \pa{t} \mathcal{R} + 8 \gamma^\rho{}_{t \rho} &= 0,
  \label{eq1}
  \\
  \pa{\rho} \mathcal{R} - 4 \mathcal{R} + 16 \alpha^{-2} \gamma^\rho{}_{tt} &= 0,
  \label{eq2}
  \\
  \pa{t} \big( \Delta_x e^{2\rho} \big) + 4 \beta^{-2} e^{2\rho} \gamma^\rho{}_{t \rho} &= 0,
  \label{eq3}
\end{align}
and
\begin{dmath}
  \pa{\rho} \big( \Delta_x e^{2\rho} \big) + 8 - 4 \Delta_x e^{2\rho} - 16 \beta^{-2} e^{2\rho} - 4 \beta^{-2} e^{2\rho} \gamma^{\rho}{}_{\rho\rho}  + \mathcal{R} e^{2\rho} = 0.
  \label{eq4}
\end{dmath}
From equation~\eqref{eq1} it follows that,
%% can be rewritten as
%% \begin{equation}
%%   \pa{t} \big( \mathcal{R} - 8 \beta^{-2} \big) = 0,
%%   \label{eq1a}
%% \end{equation}
%% or 
\begin{equation}
  \mathcal{R} - 8 \beta^{-2} = G_1(\rho).
  \label{eq1b}
\end{equation}
In addition, since
\begin{equation}
  \Delta_x e^{2\rho}
  = e^{2\rho} \bigg[ 4 \beta^{-2} + \frac{2}{\alpha \beta} \pa{\rho} \bigg( \frac{\alpha}{\beta} \bigg) \bigg]
  \equiv  e^{2\rho} \Big[ 4 \beta^{-2} + \mathcal{F} \Big],
\end{equation}
it follows from Eqs.~\eqref{eq3} that,
%% \begin{equation}
%%   \pa{t} \big( F + 2 \beta^{-2} \big) = 0,
%%   \label{eq3a}
%% \end{equation}
%% or 
\begin{equation}
  \mathcal{F} + 2 \beta^{-2} = G_2(\rho).
  \label{eq3b}
\end{equation}

Equations~\eqref{eq1b} and~\eqref{eq3b} implies that all temporal dependence on time within $\mathcal{R}$, $\mathcal{F}$ and $\beta^{-2}$, enters through an additive common function, say $B(t,\rho)$. However, the consistency of these equations with the value of $\mathcal{R}$ calculated from the metric ansatz, in Eq.~\eqref{metr2d}, sets this additive function to zero, $B(t,\rho)$. Therefore, neither $\mathcal{R}$, $\mathcal{F}$ nor $\beta^{-2}$ depend on $t$. By extension---from the definition of $\mathcal{F}$---, $\alpha$ could depend on time through a multiplicative function. Nonetheless, this multiplicative function can be eliminated of the problem by a redefinition of the \emph{time} coordinate.

At this stage, we have proved that \emph{any spherically symmetric (metric) solution of the Eqs.~\eqref{SimpleEOM} is necessarily static}, which is one of the main statements of Birkhoff's theorem.

This result, simplifies the Eq.~\eqref{hugeeq} but the equations are still non-linear. However, one can get ride off the non-linearity by setting $A(r) = B(r)$. In this \emph{linearized} regime, according to Eq.~\eqref{simplef} the general (metric) solution is the (Anti-)de-Sitter--Schwarzschild metric.

Although we restricted ourselves to the \emph{linearized} regime of the field equations, it is possible to go a bit further in the nonlinear regime. As a preliminary result~\cite{OCF-future2}, we have found that using the Frobenius methods with a Laurent series expansion, the leading order of the solutions is three-fold in the radial coordinate. These solutions correspond to the pairs given by $(\alpha,\beta)$---$\alpha$ (resp. $\beta)$ is the leading order of the series expansion for the function $A(r)$ (resp. $B(r)$)---equals to $(-1,-1)$, $(2,-2)$ and $(0,-1)$.
%%Two of these branches coincide with the reciprocal and quadratic dependence on $r$ of $A$ and $B$, while the third one gives novel dependence,
%% \begin{equation}
%%   A(r) = 1 + \frac{\alpha}{r} + \beta r^2 + \frac{\gamma}{r^2},
%% \end{equation}
%% and
%% \begin{equation}
%%   B(r) = 1 + \frac{\alpha}{r} + (\beta + \gamma) r^2.
%% \end{equation}

%% \section{extra... DELETE IT}

%% \begin{widetext}
%%   \begin{dmath*}
%%     todo2 :=
%%     %% \Lambda {\epsilon}_{\alpha \beta \gamma \delta} (B_1 + \frac{1}{3}\, B_2) {t_{\perp}}_{\eta}\,^{\alpha \beta} {t_{\perp}}^{\eta \gamma \delta} + \Lambda {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}_{\eta}{{ta}^{\alpha \beta}}\,  (4\, B_1 + \frac{4}{3}\, B_2) {t_{\perp}}^{\eta \gamma \delta} + \Lambda {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}^{\alpha}\,_{\eta}{{t_{\mathtt{v}}}^{\beta}}\,  (4\, B_1 + \frac{4}{3}\, B_2) {t_{\perp}}^{\eta \gamma \delta} + \Lambda {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}_{\eta}{{ta}^{\alpha \beta}}\,  {\nabla}^{\eta}{{ta}^{\gamma \delta}}\,  (4\, B_1 + \frac{4}{3}\, B_2) + \Lambda {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}_{\eta}{{ta}^{\alpha \beta}}\,  {\nabla}^{\gamma}{{ts}^{\delta \eta}}\,  (8\, B_1 + \frac{8}{3}\, B_2) + \Lambda {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}_{\eta}{{ta}^{\alpha \beta}}\,  {\nabla}^{\gamma \eta}{{t_{\mathtt{v}}}^{\delta}}\,  (8\, B_1 + \frac{8}{3}\, B_2) + \Lambda {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}^{\alpha}\,_{\eta}{{t_{\mathtt{v}}}^{\beta}}\,  {\nabla}^{\gamma \eta}{{t_{\mathtt{v}}}^{\delta}}\,  (4\, B_1 + \frac{4}{3}\, B_2)
%%     + \Lambda ( - B_3 + \frac{2}{3}\, B_4) {at}_{\alpha} {t_{\perp}}_{\beta}\,^{\alpha \beta} + \Lambda {\nabla}^{\alpha}\,_{\beta}{{t_{\mathtt{v}}}^{\beta}}\,  ( - B_3 + \frac{2}{3}\, B_4) {at}_{\alpha} + \Lambda {\nabla}^{\alpha}\,_{\alpha}{{t_{\mathtt{v}}}^{\beta}}\,  (B_3 - \frac{2}{3}\, B_4) {at}_{\beta} + \Lambda {\nabla}_{\alpha}{aL}\,  ( - B_3 + \frac{2}{3}\, B_4) {t_{\perp}}_{\beta}\,^{\alpha \beta} + \Lambda {\nabla}_{\alpha}{aL}\,  {\nabla}^{\alpha}\,_{\beta}{{t_{\mathtt{v}}}^{\beta}}\,  ( - B_3 + \frac{2}{3}\, B_4) + \Lambda {\nabla}_{\alpha}{aL}\,  {\nabla}^{\beta}\,_{\beta}{{t_{\mathtt{v}}}^{\alpha}}\,  (B_3 - \frac{2}{3}\, B_4) + C_1 \Lambda {gt}_{\alpha \beta \gamma} {t_{\perp}}^{\beta \alpha \gamma} + C_1 \Lambda {\nabla}_{\alpha}{{gs}_{\beta \gamma}}\,  {t_{\perp}}^{\beta \alpha \gamma} + 3\, C_1 \Lambda {\nabla}_{\alpha}{{gs}_{\beta \gamma}}\,  {\nabla}^{\beta}{{ta}^{\alpha \gamma}}\,  - C_1 \Lambda {\nabla}_{\alpha}{{gs}_{\beta \gamma}}\,  {\nabla}^{\beta}{{ts}^{\alpha \gamma}}\,  - C_1 \Lambda {\nabla}^{\alpha}\,_{\alpha}{{ts}^{\beta \gamma}}\,  {gs}_{\beta \gamma} - C_1 \Lambda {\nabla}_{\alpha}{{gs}_{\beta \gamma}}\,  {\nabla}^{\beta \gamma}{{t_{\mathtt{v}}}^{\alpha}}\, %
%%     + C_1 \Lambda {\nabla}_{\alpha}{{gs}_{\beta \gamma}}\,  {\nabla}^{\alpha \beta}{{t_{\mathtt{v}}}^{\gamma}}\,  + C_1 \Lambda {\nabla}_{\alpha \beta}{{gv}_{\gamma}}\,  {t_{\perp}}^{\alpha \beta \gamma} + 2\, C_1 \Lambda {\nabla}_{\alpha}{{ts}^{\beta \gamma}}\,  {\nabla}^{\alpha}\,_{\beta}{{gv}_{\gamma}}\,  - C_1 \Lambda {\nabla}_{\alpha \beta}{{gv}_{\gamma}}\,  {\nabla}^{\alpha \gamma}{{t_{\mathtt{v}}}^{\beta}}\,  + C_1 \Lambda {\nabla}_{\alpha \beta}{{gv}_{\gamma}}\,  {\nabla}^{\alpha \beta}{{t_{\mathtt{v}}}^{\gamma}}\,  + 2\, C_1 \Lambda {\nabla}_{\alpha}{{ts}^{\beta \gamma}}\,  {\nabla}_{\beta \gamma}{{gvp}^{\alpha}}\,  + 2\, C_1 \Lambda {\nabla}_{\alpha}{{ts}^{\beta \gamma}}\,  {\nabla}^{\alpha}\,_{\beta \gamma}{gL}\,  + 2\, C_1 \Lambda {\nabla}^{\alpha}\,_{\beta}{{t_{\mathtt{v}}}^{\gamma}}\,  {\nabla}_{\alpha}\,^{\beta}\,_{\gamma}{gL}\,  + 3\, C_1 {\nabla}^{\alpha}\,_{\alpha}{{gs}_{\beta \gamma}}\,  {\nabla}^{\beta}\,_{\delta}{{ta}^{\gamma \delta}}\,  - C_1 {\nabla}^{\alpha}\,_{\alpha}{{gs}_{\beta \gamma}}\,  {\nabla}^{\delta}\,_{\delta}{{ts}^{\beta \gamma}}\,  - C_1 {\nabla}^{\alpha}\,_{\alpha}{{gs}_{\beta \gamma}}\,  {\nabla}^{\beta \delta}\,_{\delta}{{t_{\mathtt{v}}}^{\gamma}}\,  + 3\, C_1 {\nabla}_{\alpha \beta}{{gp}^{\alpha}\,_{\gamma}}\,  {\nabla}^{\beta}\,_{\delta}{{ta}^{\gamma \delta}}\,  + 3\, C_1 {\nabla}_{\alpha \beta}{{gp}^{\alpha}\,_{\gamma}}\,  {\nabla}^{\gamma}\,_{\delta}{{ta}^{\beta \delta}}\,  - 2\, C_1 {\nabla}_{\alpha \beta}{{gp}^{\alpha}\,_{\gamma}}\,  {\nabla}^{\delta}\,_{\delta}{{ts}^{\beta \gamma}}\,  - C_1 {\nabla}_{\alpha \beta}{{gp}^{\alpha}\,_{\gamma}}\,  {\nabla}^{\beta \delta}\,_{\delta}{{t_{\mathtt{v}}}^{\gamma}}\,  - C_1 {\nabla}_{\alpha \beta}{{gp}^{\alpha}\,_{\gamma}}\,  {\nabla}^{\gamma \delta}\,_{\delta}{{t_{\mathtt{v}}}^{\beta}}\,  + {\nabla}^{\alpha}\,_{\beta}{{ta}^{\beta \gamma}}\,  {\nabla}_{\alpha}\,^{\delta}\,_{\delta}{{gv}_{\gamma}}\,  ( - 2\, C_1 + 2\, C_2) + {\nabla}^{\alpha}\,_{\beta}{{ta}^{\beta \gamma}}\,  {\nabla}_{\gamma}\,^{\delta}\,_{\delta}{{gv}_{\alpha}}\,  ( - C_1 - 2\, C_2) + {\nabla}^{\alpha}\,_{\alpha}{{ta}^{\beta \gamma}}\,  {\nabla}_{\beta}\,^{\delta}\,_{\delta}{{gv}_{\gamma}}\,  ( - C_1 + 4\, C_2) - C_1 {\nabla}^{\alpha}\,_{\alpha}{{ts}^{\beta \gamma}}\,  {\nabla}_{\beta}\,^{\delta}\,_{\delta}{{gv}_{\gamma}}\, %
%%     + {\nabla}^{\alpha}\,_{\alpha \beta}{{gv}_{\gamma}}\,  {\nabla}^{\beta \delta}\,_{\delta}{{t_{\mathtt{v}}}^{\gamma}}\,  ( - C_1 + 2\, C_2) + {\nabla}_{\alpha}{{gt}^{\beta}\,_{\beta \gamma}}\,  {\nabla}^{\gamma}\,_{\delta}{{ta}^{\alpha \delta}}\,  ( - 2\, C_1 + 2\, C_2) + {\nabla}_{\alpha}{{gt}^{\beta}\,_{\beta \gamma}}\,  {\nabla}^{\alpha}\,_{\delta}{{ta}^{\gamma \delta}}\,  ( - C_1 - 2\, C_2) + {\nabla}_{\alpha}{{gt}^{\beta}\,_{\beta \gamma}}\,  {\nabla}^{\delta}\,_{\delta}{{ta}^{\alpha \gamma}}\,  ( - C_1 + 4\, C_2) + C_1 {\nabla}_{\alpha}{{gt}^{\beta}\,_{\beta \gamma}}\,  {\nabla}^{\delta}\,_{\delta}{{ts}^{\alpha \gamma}}\,  + {\nabla}_{\alpha}{{gt}^{\beta}\,_{\beta \gamma}}\,  {\nabla}^{\gamma \delta}\,_{\delta}{{t_{\mathtt{v}}}^{\alpha}}\,  (C_1 - 2\, C_2) - 3\, C_1 {\nabla}_{\alpha \beta}{{gp}^{\gamma}\,_{\gamma}}\,  {\nabla}^{\alpha}\,_{\delta}{{ta}^{\beta \delta}}\,  + C_1 {\nabla}_{\alpha \beta}{{gp}^{\gamma}\,_{\gamma}}\,  {\nabla}^{\delta}\,_{\delta}{{ts}^{\alpha \beta}}\,  + C_1 {\nabla}_{\alpha \beta}{{gp}^{\gamma}\,_{\gamma}}\,  {\nabla}^{\alpha \delta}\,_{\delta}{{t_{\mathtt{v}}}^{\beta}}\,  + 3\, C_1 {\nabla}^{\alpha}\,_{\beta}{{ta}^{\beta \gamma}}\,  {\nabla}_{\alpha \gamma}\,^{\delta}{{gv}_{\delta}}\,  + C_1 {\nabla}^{\alpha}\,_{\alpha}{{ts}^{\beta \gamma}}\,  {\nabla}_{\beta \gamma}\,^{\delta}{{gv}_{\delta}}\,  + C_1 {\nabla}^{\alpha}\,_{\beta \gamma}{{gv}_{\alpha}}\,  {\nabla}^{\beta \delta}\,_{\delta}{{t_{\mathtt{v}}}^{\gamma}}\,  + 2\, C_2 {\nabla}_{\alpha}{{gt}^{\beta}\,_{\beta \gamma}}\,  {\nabla}^{\alpha \delta}\,_{\delta}{{t_{\mathtt{v}}}^{\gamma}}\,  - 2\, C_2 {\nabla}^{\alpha}\,_{\alpha \beta}{{gv}_{\gamma}}\,  {\nabla}^{\gamma \delta}\,_{\delta}{{t_{\mathtt{v}}}^{\beta}}\,
%%     %% + 4\, E_1 {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}^{\eta}\,_{\eta}{{ta}^{\alpha \beta}}\,  {\nabla}^{\lambda}\,_{\lambda}{{ta}^{\gamma \delta}}\,  - 8\, E_1 {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}^{\alpha}\,_{\eta}{{ta}^{\beta \eta}}\,  {\nabla}^{\lambda}\,_{\lambda}{{ta}^{\gamma \delta}}\,  + 8\, E_1 {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}^{\eta}\,_{\eta}{{ta}^{\alpha \beta}}\,  {\nabla}^{\gamma \lambda}\,_{\lambda}{{t_{\mathtt{v}}}^{\delta}}\,  + 4\, E_1 {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}^{\alpha}\,_{\eta}{{ta}^{\beta \eta}}\,  {\nabla}^{\gamma}\,_{\lambda}{{ta}^{\delta \lambda}}\,  - 8\, E_1 {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}^{\alpha}\,_{\eta}{{ta}^{\beta \eta}}\,  {\nabla}^{\gamma \lambda}\,_{\lambda}{{t_{\mathtt{v}}}^{\delta}}\,  + 4\, E_1 {\epsilon}_{\alpha \beta \gamma \delta} {\nabla}^{\alpha \eta}\,_{\eta}{{t_{\mathtt{v}}}^{\beta}}\,  {\nabla}^{\gamma \lambda}\,_{\lambda}{{t_{\mathtt{v}}}^{\delta}}\, %
%%     + 3\, E_2 {\nabla}_{\alpha}{{at}_{\beta}}\,  {\nabla}^{\gamma}\,_{\gamma}{{ta}^{\alpha \beta}}\,  - 3\, E_2 {\nabla}_{\alpha}{{at}_{\beta}}\,  {\nabla}^{\alpha}\,_{\gamma}{{ta}^{\beta \gamma}}\,  + 2\, E_2 {\nabla}_{\alpha}{{at}_{\beta}}\,  {\nabla}^{\alpha \gamma}\,_{\gamma}{{t_{\mathtt{v}}}^{\beta}}\,  - E_2 {\nabla}_{\alpha}{{at}_{\beta}}\,  {\nabla}^{\beta \gamma}\,_{\gamma}{{t_{\mathtt{v}}}^{\alpha}}\,  - 3\, E_2 {\nabla}_{\alpha \beta}{aL}\,  {\nabla}^{\alpha}\,_{\gamma}{{ta}^{\beta \gamma}}\,  + E_2 {\nabla}_{\alpha \beta}{aL}\,  {\nabla}^{\alpha \gamma}\,_{\gamma}{{t_{\mathtt{v}}}^{\beta}}\,  + E_2 {\nabla}_{\alpha}{{at}_{\beta}}\,  {\nabla}^{\gamma}\,_{\gamma}{{ts}^{\alpha \beta}}\,  + E_2 {\nabla}_{\alpha \beta}{aL}\,  {\nabla}^{\gamma}\,_{\gamma}{{ts}^{\alpha \beta}}\, ;
%%   \end{dmath*}
%% \end{widetext}